Spin canonical divisors of log stacky curves

### 2015-06-23 stacky_write_up/spin_cring.tex
finish base/exception cases, add larger pic

\usepackage{amsmath}
\usepackage{amsxtra}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage[all]{xy}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{mathrsfs}
\usepackage{subfiles}
\usepackage{mathpazo}
\usepackage{euler}
\usepackage{hyperref}
\usepackage{color}
\usepackage{longtable}

@@

\todo{Move this lemma:}
\begin{lem}\label{lem:add_point_effective_divisor_no_deg_1_generators}
Let $(X,\Delta,\halfcan)\to (X,\Delta',\halfcan)$ be the natural map of genus $g$ log curves with $\Delta=\Delta'+P$, such that $2\halfcan\sim K_X+\Delta$, $2\halfcan'\sim K_X+\Delta'$, and $\Delta=\Delta'+P$ with $\Delta'\ne 0$ for some $P\not\in supp(\halfcan)$.  
Suppose $\deg(\lfloor{\halfcan}\rfloor)=\deg(\lfloor \halfcan'\rfloor)\ge 0$, so $R_{\halfcan}$ has a
generator $u$ in degree 1, and $R_{\halfcan'}$ has no new generators in degree 0.  Suppose $D$ is generated by $u, x_1, \ldots, x_m$ in degree at most $\bar{d}$ with relations generated 
in degree at most $\bar{\tau}$.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1, y_2, y_3$ in degrees 2, 2, and 3 respectively.  
\item If $I$ and $I'$ are the ideal of relations of $R_D$ and $R_{D'}$ respectively, then

\begin{align*}
	\initial_\prec(I') &= \initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 &+ \langle y_i x_j \rangle \\
										 &+ \langle y_3^2, y_3y_2, y_2y_1, y_1^2\rangle \\
\end{align*}
\item A Grobner basis with initial terms described above minimally generates $I'$ over $I$.
\end{itemize}
\end{lem}
\begin{proof}
Similar to all the other lemmas\todo{combine sections?}
\end{proof}

\begin{lem}\label{lem:add_point_effective_divisor_deg_1_but_not_deg_2_generators}
Let $2\halfcan=K_X$ of a non-stacky curve $X$ such that $h^0(X, \halfcan) > 0$.  Then for any point $P$, 
\[
	H^0(X, \halfcan+P)=H^0(X,\halfcan).
\]
\end{lem}
\begin{proof}
\todo{add proof}
\end{proof}

@@

Therefore, the canonical ring $R'$ has presentation $R' =
k[x_{7, 1}, x_{7, 2}, x{5, 1}, x_{3, 1}] / I$ with
initial ideal
\[
	in_\prec(I) = < x_{7, 1} x_{3, 1}, x_{7, 1}^2 >
\]

\noindent
generated by quadratics under grevlex with $x_{i, j} \prec x_{k, l}$
if $i > j$ or $i = j$ and $j < l$. These quadratics must also 
determine minimal generators for $I$ by Lemma
~\ref{lem:minimal_quadratic}. Thus, $R'$ is generated up to degree
$e = 7$ with relations up to degree $2e = 14$, as desired.
We can also see that $e_i > \deg z$ for any generator $z$ of $R'$
since $e_i = 9 > e$ for all $i \in J$.

\todo{show only relations}

Admissibility is the remaining condition to check. WLOG, we can
let $x_{7, 1}$ correspond to the generator with maximal pole order
at $P_2$ and $x_{7, 2}$ correspond to the generator with maximal
pole order at $P_3$. Use the presentation given above with $y
_{2, e_2'} := x_{7, 1}$ and $y_{3, e_3'} :=  x_{7, 2}$. We see that
it immediately satisfies (Ad-i) of Definition ~\ref{defn:admissible}.
It also satisfies (Ad-ii) because $y_{i, e_i'}$ is the only generator
lying on the line $-ord_{P_i}(z) = \deg z \frac{3d}{7}$ in the
$(\deg z, -ord_{P_i}(z))$ lattice and the other generators
must lie below the line as seen below:

\begin{center}
\includegraphics{pics/spin-377-pic-pics.pdf}
\end{center}

@@

Furthermore, each case satisfies (Ad-iii) as demonstrated below.
Notice that the $e_i$ and $\{e_j' : j \neq i\}$ are equivalent
for any choice of $i \in J$ for these cases, so $\deg \lfloor e_i L
\rfloor$ and $\max_{d \geq 0} \#S_{(\sigma, J)}(i)$ are
independent of the choice of $i$.

@@

All that remains to be shown is that these cases satisfy
(Ad-i) and (Ad-ii) and that $in_\prec(I')$ is minimally generated by
products of two monomials. The procedure for verifying these
are similar to those for the given example. Below, we
give a table for the explicit generators and initial ideals.
\todo{make table, move this before (Ad-iii)?}
\end{proof}

@@

\begin{align*}
	S &:= \{(0; 3, 3, k; 0) : 3 \leq k \leq 9 \text{ odd}\} \\
		&\cup \{(0; 3, b, c; 0) : 5 \leq b \leq c \leq 7\} \\
		&\cup \{(0; 5, 5, 5; 0), (0; 3, 3, 3, 3; 0)\}
\end{align*}

@@

	$(0; 3, 3, 3, 3; 0)$ & $\{3, 3, 4\}$ & $\{12\}$ & $3$ \\	\hline

	$(0; 3, 7, 7; 0)$ & $\{3, 5, 7, 7\}$ & $\{10, 14\}$ & $7$ \\	\hline
\end{longtable}

@@

\begin{rem}
These cases, apart from $(0; 3, 7, 7; 0)$, give all of the
exceptions to the $e$ and $2e$ bounds on the generator and relation
degree. Notice that each of these exceptional cases, apart from
$(0; 3, 5, 7; 0)$, corresponds to a signature with non-generic
saturation (i.e. saturation not equal to 3, 5, or 9 as seen
in the table in Subsection ~\ref{ssec:g_0_saturation}). Exceptional
saturation can be viewed as ``forcing'' generators (and thus
relations) in higher degrees than in the generic case.
\end{rem}

\ssec{Application of induction}
\label{ssec:g_0_noneff_induction}
Now we can combine this.

@@

\section{Further Research}
\begin{itemize}
	\item More on fractional modular forms
	\item Make explicit generators and relations for base case of non-stacky higher genus half-canonical divisors
	\item In the case $g \geq 2$, can the bound from Proposition ~\ref{prop:relation_generation} be reduced from 11 to 10 (which would then be sharp by Remark ~\ref{rem:relations_generation_ten})? \todo{Possibly think more about this} 
	\item Generic Initial ideal (and other genericness properties?)
\end{itemize}

>>>>

\usepackage{amsmath}
\usepackage{amsxtra}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage[all]{xy}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{mathrsfs}
\usepackage{subfiles}
\usepackage{mathpazo}
\usepackage{euler}
\usepackage{hyperref}
\usepackage{color}
\usepackage{longtable}
\usepackage{float}

@@

\todo{Move this lemma:}
\begin{lem}\label{lem:add_point_effective_divisor_no_deg_1_generators}
Let $(X,\Delta,\halfcan)\to (X,\Delta',\halfcan)$ be the natural map of genus $g$ log curves with $\Delta=\Delta'+P$, such that $2\halfcan\sim K_X+\Delta$, $2\halfcan'\sim K_X+\Delta'$, and $\Delta=\Delta'+P$ with $\Delta'\ne 0$ for some $P\not\in supp(\halfcan)$.  
Suppose $\deg(\lfloor{\halfcan}\rfloor)=\deg(\lfloor \halfcan'\rfloor)\ge 0$, so $R_{\halfcan}$ has a
generator $u$ in degree 1, and $R_{\halfcan'}$ has no new generators in degree 0.  Suppose $D$ is generated by $
u, x_1, \ldots, x_m$ in degree at most $\bar{d}$ with relations generated 
in degree at most $\tau $.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1,...,y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

@@

Let $I$ be the ideal generated by these relations in
$k[x_{7, 1}, x_{7, 2}, x{5, 1}, x_{3, 1}]$. Under grlex with
$x_{i, j} \prec x_{k, l}$ if $i > j$ or $i = j$ and $j < l$,
the initial ideal of $I$ is
\[
	in_\prec(I) = < x_{5, 1}^2, x_{7, 1}^2 >
\]

To demonstrate that these two relations generate all 
relations, we show that $\dim (R')_d = \dim
(k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / I)_d$ for all degrees
$d \geq 0$. In particular,

\begin{align*}
	\dim (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / I)_d &=
	\dim (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}]	/ in_\prec(I))_d \\
	&= \dim (k[x_{3, 1}, x_{5, 1}, x_{7, 1}, x_{7, 2}] / < x_{5, 1}^2,
	x_{7, 1}^2 >
\end{align*}

\noindent
for all $d \geq 0$. But the fact $\dim (R')_d = \dim (k[x_{3, 1},
x_{5, 1}, x_{7, 1}, x_{7, 2}] / < x_{5, 1}^2,	x_{7, 1}^2 >$ can
be checked by looking at the $k$-basis at each degree $d$.

Therefore, the canonical ring $R'$ has presentation $R' =
k[x_{7, 1}, x_{7, 2}, x_{5, 1}, x_{3, 1}] / I$ with
initial ideal $in_\prec(I)$ generated by quadratics under grlex
with $x_{i, j} \prec x_{k, l}$ if $i > j$ or $i = j$ and $j < l$.
These quadratics must also determine minimal generators for $I$ by
Lemma ~\ref{lem:minimal_quadratic}. Thus, $R'$ is generated up to
degree $e = 7$ with relations up to degree $2e = 14$, as desired.
We can also see that $e_i > \deg z$ for any generator $z$ of $R'$
since $e_i = 9 > e$ for all $i \in J$.

Admissibility is the remaining condition to check. We can assume
that $x_{7, 1}$ corresponds to the generator with maximal pole order
at $P_2$ and $x_{7, 2}$ correspond to the generator with maximal
pole order at $P_3$ (relabel otherwise). Use the presentation given
above with $y _{2, e_2'} := x_{7, 1}$ and $y_{3, e_3'} :=  x_{7, 2}$.
We see that it immediately satisfies (Ad-i) of Definition
~\ref{defn:admissible}. We may also choose pole orders of the generators
such that $y_{i, e_i'}$ is the only generator lying on the line
$-ord_{P_i}(z) = \deg z \frac{3d}{7}$ in the $(\deg z, -ord_{P_i}(z))$
 lattice and with the other generators lying below the line as seen
in Figure ~\ref{fig:377} (e.g. the pole orders $(-ord_{P_1}(z)$,
$-ord_{P_2}(z)$, $-ord_{P_3}(z))$ may be chosen to be $(1, 1, 1)$,
$(1, 2, 2)$, $(2, 3, 2)$, and $(2, 2, 3)$ for $x_{3, 1}, x_{5, 1},
x_{7, 1}, x_{7, 2}$ respectively).

\begin{figure}[H]
\label{fig:377}
\includegraphics{pics/spin-377-pic-pics.pdf} \\
\caption{Generators in the $(\deg z, -ord_{P_i}(z))$ lattice}
\end{figure}

@@

We can also always find a presentation for these cases such that
they satisfy (Ad-i) and (Ad-ii) and that $in_\prec(I')$ is
minimally generated by products of two monomials. The procedure for
verifying these are similar to those for the given example.

Furthermore, each case always satisfies (Ad-iii) as demonstrated below.
Notice that the $e_i$ and $\{e_j' : j \neq i\}$ are equivalent
for any choice of $i \in J$ for these cases, so $\deg \lfloor e_i L
\rfloor$ and $\max_{d \geq 0} \#S_{(\sigma, J)}(i)$ are
independent of the choice of $i$.

@@

Thus, all of the cases are admissible and satisfy the additional
desired conditions.
\end{proof}

@@

\begin{align*}
	S &:= \{(0; 3, 3, k; 0) : 3 \leq k \leq 9 \text{ odd}\} \\
		&\cup \{(0; 3, 5, 5; 0) ,(0; 3, 5, 7; 0), (0; 5, 5, 5; 0), (0; 3, 3, 3, 3; 0)\}
\end{align*}

@@

	$(0; 3, 3, 3, 3; 0)$ & $\{3, 3, 4\}$ & $\{12\}$ & $3$ \\	\hline
\end{longtable}

@@

\begin{rem}
These cases give all of the exceptions to the $e$ and $2e$ bounds on
the generator and relation degree. Notice that each of these
exceptional cases, apart from $(0; 3, 5, 7; 0)$, corresponds to a
signature with non-generic saturation (i.e. saturation not equal to
3, 5, or 9 as seen in the table in Subsection
~\ref{ssec:g_0_saturation}). Exceptional saturation can be viewed as
``forcing'' generators (and thus relations) in higher degrees than
in the generic case.
\end{rem}

\ssec{Application of induction}
\label{ssec:g_0_noneff_induction}
Now we can combine this.
\todo{write this section}

@@

\section{Further Research}
\begin{itemize}
	\item More on fractional modular forms
	\item Make explicit generators and relations for base case of non-stacky higher genus half-canonical divisors
	\item In the case $g \geq 2$, can the bound from Proposition ~\ref{prop:relation_generation} be reduced from 11 to 10 (which would then be sharp by Remark ~\ref{rem:relations_generation_ten})? \todo{Possibly think more about this} 
\end{itemize}
